Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-07-18 supplement.tex
supp results for nama map

\begin{itemize}
    \item Using YRI and CEU, inferred parameters for each of our highlighted models
        did not change (all well within one SD of the inferred parameters of the
        original fit models)
    \item TODO: run fits with TSI
    \item Fits using the Nama recombination map, even though there were slight
        differences in some of the LD statistics (primarily a few noticeable
        differences in the $\pi_2$ statistics), also resulted in model fits that
        were very similar to the original fit models.
    \item Fits with alternate thresholds of Nama-inferred ancestry: TODO:
        fit continuous migration and the merger with stem migration model
\end{itemize}

@@

Because CEU and GBR are genetically closely related, we also tested the
consistency of inferred parameters when replacing GBR with Toscani from Italy,
a more distantly related European population set than CEU. \aprcomment{TODO}

@@

@@

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-omni-nama-comparison.pdf}
    \caption{
        \textbf{Comparison of LD statistics parsed using OMNI YRI and
        Nama-specific recombination map.}
        The OMNI YRI recombination map and the Nama-inferred map
        \citep{Van_Eeden2021-od} differ in two primary ways: 1) the YRI and
        Nama have relatively highly diverged ancestries, and 2) the YRI
        map was inferred from low-density array data, while the Nama map
        was inferred from whole-genome sequencing data.
        Nonetheless, LD statistics computed under recombination rates from
        the two maps remain largely concordant, and inferences are not
        strongly affected by recombination map choice (Section~\ref{sec:supp-results}).
        Shading represents 95\% confidence intervals from the Nama-inferred
        map. Notations for each statistic are described in
        section~\ref{sec:supp-ld-stats} and indexed by populations (0: Nama, 1:
        Mende, 2: Gumuz, 3: Amhara/Oromo, 4: British).
    }
    \label{fig:supp-omni-hapmap-comparison}
\end{figure}

>>>>

@@

Because CEU and GBR are genetically closely related, we also tested the
consistency of inferred parameters when replacing GBR with Toscani from Italy,
a more distantly related European population set than GBR. Most parameters in
each of the tested models were consistent with parameters inferred using GBR.
The only large difference was found in fitting the continuous migration model,
Nama divergence was $\approx200$ka, rather than $135$ka. Migration rates during
this early period of the model also shifted to larger values between the stem
populations, demonstrating that divergence times and migration rates are
confounded parameters in this model. Notably, the multiple merger model (which
provides a better fit to the data) was inferred to have consistent divergence
and merger times between the models fit using GBR and TSI, with ancestral
populations ``reshuffling'' between 100 and 120ka.

@@

The LD statistics computed using the OMNI YRI recombination map and the
Nama-specific recombination map were very similar, with only a handful of
statistics showing slight visible differences
(Figure~\ref{fig:supp-omni-nama-comparison}). We therefore expected inferred
demographic models to be highly consistent when fit to the two sets of
statistics. Across each of our models, inferred parameters from fits using data
from the Nama-specific map were unchanged, with best-fit values all within one
estimated standard error from the inferred parameters from the OMNI YRI map.

Among human populations, the Nama are relatively highly diverged from the
sampled populations used in estimating the OMNI recombination maps (such as the
YRI), and the two recombination maps were inferred using different
computational methods from different types of sequencing data (arrays vs. whole
genomes). Therefore, differences between these two maps are likely near the
upper bound for differences in recombination rates between present-day human
populations \citep{Van_Eeden2021-od}. Because our fits were highly consistent
when using the two recombination maps, we conclude that inferences using the
diversity and LD statistics here are robust to variation in recombination rates
between present-day populations as well as to any differences in inferred
recombination rates due to true recombination rate differences or typical
statistical error in recombination rate estimation.

\subsubsection{Varying Nama ancestry-proportion thresholds}

@@

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-omni-nama-comparison.pdf}
    \caption{
        \textbf{Comparison of LD statistics parsed using OMNI YRI and
        Nama-specific recombination map.}
        The OMNI YRI recombination map and the Nama-inferred map
        \citep{Van_Eeden2021-od} differ in two primary ways: 1) the YRI and
        Nama have relatively highly diverged ancestries, and 2) the YRI
        map was inferred from low-density array data, while the Nama map
        was inferred from whole-genome sequencing data.
        Nonetheless, LD statistics computed under recombination rates from
        the two maps remain largely concordant, and inferences are
        unaffected by recombination map choice (Section~\ref{sec:supp-results}).
        Shading represents 95\% confidence intervals from the Nama-inferred
        map. Notations for each statistic are described in
        section~\ref{sec:supp-ld-stats}, and indexes represent populations
        (0: Nama, 1: Mende, 2: Gumuz, 3: Amhara/Oromo, 4: British).
    }
    \label{fig:supp-omni-nama-comparison}
\end{figure}
